---
title: "Code for figure 1"
description: |
  Code for figure 1.
author: 
  - name: Peng Gao
    affiliation: Stanford School of Medicine
  - name: Xiaotao Shen
    url: https://www.shenxt.info/
    affiliation: Stanford School of Medicine
date: "`r Sys.Date()`"
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = '../')
# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages

```


Current environmental health studies mainly focus on limited stressors and populations, which can not address the complex and personal nature of exposures. Here we longitudinally profiled personal exposome and internal multi-omics to investigate how the exposome shaped an individual’s phenome chronically for the first time. Thousands of chemical and biological substances were identified in the personal exposome cloud. We performed correlational analyses within the exposome and inter -omics analyses between the exposome and gut microbiome, metabolome, and proteome, respectively. Furthermore, the exposome and multi-omics correlations were cross-validated using clinical test results. Our results showed agrochemicals and fungi predominated in the highly diverse and dynamic exposome, and the individual’s immune system, kidneys, and liver played essential roles in response to the personal exposome. Overall, our findings demonstrated that the exposome interacts with internal multi-omics dynamically and could impact one’s health status.


## **Load libraries**

```{r, eval=TRUE, message=FALSE, cache=TRUE, warning=FALSE, echo=TRUE}
library(tinyTools)
library(tidyverse)
library(circlize)
library(ComplexHeatmap)
library(patchwork)
```


## **Set work directory**

```{r, eval=TRUE, message=FALSE, cache=TRUE, warning=FALSE, echo=TRUE}
tinyTools::setwd_project()
getwd()
```

## **Figure 1b**

```{r, eval=TRUE, message=FALSE, cache=TRUE, warning=FALSE, echo=TRUE}


# load data
##exposome chemical
load("data_20200511/exposome/sample_info")
load("data_20200511/exposome/expression_data")

exposomeChemical_sample_info <- sample_info
exposomeChemical_expression_data <- expression_data

exposomeChemical_expression_data <-
  exposomeChemical_expression_data %>%
  t() %>% 
  as.data.frame() %>%
  rownames_to_column(var = "Sample_ID") %>%
  mutate(Sample_ID = stringr::str_replace(Sample_ID, "_[0-9]{1,2}", "")) %>%
  plyr::dlply(.variables = "Sample_ID") %>%
  lapply(function(x) {
    apply(x[, -1], 2, mean)
  }) %>%
  do.call(rbind, .) %>% 
  t() %>% 
  as.data.frame()

##metabolome
load("data_20200511/metabolome/sample_info")
load("data_20200511/metabolome/expression_data")
metabolome_sample_info <- sample_info
metabolome_expression_data <- expression_data

##proteome
load("data_20200511/proteome/sample_info")
load("data_20200511/proteome/expression_data")
proteome_sample_info <- sample_info
proteome_expression_data <- expression_data

##microbiome
load("data_20200511/microbiome/dna_sample_info")
load("data_20200511/microbiome/dna_expression_data")
exposomeBiological_sample_info <- dna_sample_info
exposomeBiological_expression_data <- dna_expression_data
exposomeBiological_expression_data <- 
  exposomeBiological_expression_data[grep("phylum", rownames(exposomeBiological_expression_data)),]
  
##cytokine
load("data_20200511/cytokine/sample_info")
load("data_20200511/cytokine/expression_data")
cytokine_sample_info <- sample_info
cytokine_expression_data <- expression_data

##gut microbiome
load("data_20200511/gut_microbiome/sample_info")
load("data_20200511/gut_microbiome/expression_data")
gutmicrobiome_sample_info <- sample_info
gutmicrobiome_expression_data <- expression_data

gutmicrobiome_sample_info <-
  gutmicrobiome_sample_info %>% 
  dplyr::distinct(CollectionDate, .keep_all = TRUE)

gutmicrobiome_expression_data <-
  gutmicrobiome_expression_data[,gutmicrobiome_sample_info$sample_id]

##environment
load("data_20200511/environment/sample_info")
load("data_20200511/environment/expression_data")
environment_sample_info <- sample_info
environment_expression_data <- expression_data

##lab test
load("data_20200511/lab_test/sample_info")
load("data_20200511/lab_test/expression_data")
lab_sample_info <- sample_info
lab_expression_data <- expression_data

###begin circlize

##exposome chemical
exposomeChemical_sample_info1 <- 
  exposomeChemical_sample_info %>% 
  dplyr::select(date = start_date) %>% 
  dplyr::mutate(date = as.Date(date),
                class = "exp")

####exposome toxins and carcinogens
toxin_sample_info1 <- 
  metabolome_sample_info %>% 
  dplyr::select(date = CollectionDate) %>% 
  dplyr::mutate(date = as.Date(date),
                class = "toxin")

##metabolome
metabolome_sample_info1 <- 
  metabolome_sample_info %>% 
  dplyr::select(date = CollectionDate) %>% 
  dplyr::mutate(date = as.Date(date),
                class = "met")

##proteome
proteome_sample_info1 <- 
  proteome_sample_info %>% 
  dplyr::select(date = CollectionDate) %>% 
  dplyr::mutate(date = as.Date(date),
                class = "pro")

##microbiome
exposomeBiological_sample_info1 <- 
  exposomeBiological_sample_info %>% 
  dplyr::select(date = date.start) %>% 
  dplyr::mutate(date = as.Date(date),
                class = "micro")

##cytokine
cytokine_sample_info1 <- 
  cytokine_sample_info %>% 
  dplyr::select(date = CollectionDate) %>% 
  dplyr::mutate(date = as.Date(date),
                class = "cytokine")

##gut microbiome
gutmicrobiome_sample_info1 <- 
  gutmicrobiome_sample_info %>% 
  dplyr::select(date = CollectionDate) %>% 
  dplyr::mutate(date = as.Date(date),
                class = "gutmicrobiome")

##environment
environment_sample_info1 <- 
  environment_sample_info %>% 
  dplyr::select(date = date.start) %>% 
  dplyr::mutate(date = as.Date(date),
                class = "environment")

##lab test
lab_sample_info1 <- 
  lab_sample_info %>% 
  dplyr::select(date = date.start) %>% 
  dplyr::mutate(date = as.Date(date),
                class = "lab")

all_date <- 
  data.frame(date = unique(c(
    exposomeChemical_sample_info1$date,
    metabolome_sample_info1$date,
    toxin_sample_info1$date,
    proteome_sample_info1$date,
    exposomeBiological_sample_info1$date,
    cytokine_sample_info1$date,
    gutmicrobiome_sample_info1$date,
    environment_sample_info1$date,
    lab_sample_info1$date
  ))) %>% 
  dplyr::filter(date >= "2016-01-12" & date <= "2016-03-03")
  
exposomeChemical_sample_info1 <-
  all_date %>%
  dplyr::left_join(exposomeChemical_sample_info1, by = c("date")) %>%
  dplyr::mutate(value =
                  case_when(!is.na(class) ~ "Yes",
                            TRUE ~ "No")) %>%
  dplyr::mutate(class = class[!is.na(class)][1])

toxin_sample_info1 <-
  all_date %>% 
  dplyr::left_join(toxin_sample_info1, by = c("date")) %>% 
  dplyr::mutate(value = 
                  case_when(
                    !is.na(class) ~ "Yes",
                    TRUE ~ "No"
                  )) %>% 
  dplyr::mutate(class = class[!is.na(class)][1])

metabolome_sample_info1 <-
  all_date %>% 
  dplyr::left_join(metabolome_sample_info1, by = c("date")) %>% 
  dplyr::mutate(value = 
                  case_when(
                    !is.na(class) ~ "Yes",
                    TRUE ~ "No"
                  )) %>% 
  dplyr::mutate(class = class[!is.na(class)][1])

proteome_sample_info1 <-
  all_date %>% 
  dplyr::left_join(proteome_sample_info1, by = c("date")) %>% 
  dplyr::mutate(value = 
                  case_when(
                    !is.na(class) ~ "Yes",
                    TRUE ~ "No"
                  )) %>% 
  dplyr::mutate(class = class[!is.na(class)][1])

exposomeBiological_sample_info1 <-
  all_date %>% 
  dplyr::left_join(exposomeBiological_sample_info1, by = c("date")) %>% 
  dplyr::mutate(value = 
                  case_when(
                    !is.na(class) ~ "Yes",
                    TRUE ~ "No"
                  )) %>% 
  dplyr::mutate(class = class[!is.na(class)][1])

cytokine_sample_info1 <-
  all_date %>% 
  dplyr::left_join(cytokine_sample_info1, by = c("date")) %>% 
  dplyr::mutate(value = 
                  case_when(
                    !is.na(class) ~ "Yes",
                    TRUE ~ "No"
                  )) %>% 
  dplyr::mutate(class = class[!is.na(class)][1])

gutmicrobiome_sample_info1 <-
  all_date %>% 
  dplyr::left_join(gutmicrobiome_sample_info1, by = c("date")) %>% 
  dplyr::mutate(value = 
                  case_when(
                    !is.na(class) ~ "Yes",
                    TRUE ~ "No"
                  )) %>% 
  dplyr::mutate(class = class[!is.na(class)][1])

environment_sample_info1 <-
  all_date %>% 
  dplyr::left_join(environment_sample_info1, by = c("date")) %>% 
  dplyr::mutate(value = 
                  case_when(
                    !is.na(class) ~ "Yes",
                    TRUE ~ "No"
                  )) %>% 
  dplyr::mutate(class = class[!is.na(class)][1])

lab_sample_info1 <-
  all_date %>% 
  dplyr::left_join(lab_sample_info1, by = c("date")) %>% 
  dplyr::mutate(value = 
                  case_when(
                    !is.na(class) ~ "Yes",
                    TRUE ~ "No"
                  )) %>% 
  dplyr::mutate(class = class[!is.na(class)][1])

total_sample_info <-
  rbind(
    exposomeChemical_sample_info1,
    toxin_sample_info1,
    metabolome_sample_info1,
    proteome_sample_info1,
    exposomeBiological_sample_info1,
    cytokine_sample_info1,
    gutmicrobiome_sample_info1,
    environment_sample_info1,
    lab_sample_info1
  )

value <- 
  c("Environment" = ggsci::pal_d3()(10)[1],
    "Exposome (chemical)" = ggsci::pal_d3()(10)[2],
    "Metabolome" = ggsci::pal_d3()(10)[3],
    "Proteome" = ggsci::pal_d3()(10)[4],
    "Exposome (biological)" = ggsci::pal_d3()(10)[5],
    "Gut microbiome" = ggsci::pal_d3()(10)[6],
    "Blood test" = ggsci::pal_d3()(10)[7],
    "Cytokine" = ggsci::pal_d3()(10)[8],
    "Toxins and carcinogens" = ggsci::pal_d3()(10)[9]
    )

total_sample_info <- 
  total_sample_info %>% 
  dplyr::mutate(date = as.character(date)) %>%
  dplyr::mutate(
    class = case_when(
      class == "exp" ~ "Exposome (chemical)",
      class == "micro" ~ "Exposome (biological)",
      class == "met" ~ "Metabolome",
      class == "pro" ~ "Proteome",
      class == "lab" ~ "Blood test",
      class == "cytokine" ~ "Cytokine",
      class == "gutmicrobiome" ~ "Gut microbiome",
      class == "environment" ~ "Environment",
      class == "toxin" ~ "Toxins and carcinogens"
    )
  ) %>% 
  dplyr::mutate(class = factor(class, levels = rev(
    c(
      "Exposome (chemical)",
      "Exposome (biological)",
      "Environment",
      "Metabolome",
      "Toxins and carcinogens",
      "Proteome",
      "Blood test",
      "Cytokine",
      "Gut microbiome"
    )
  )))
  

plot1 <- 
total_sample_info %>%
  ggplot(aes(date, class)) +
  geom_tile(aes(x = date, y = class, fill = value), color = "black", show.legend = FALSE) +
  scale_y_discrete(expand = expansion(mult = c(0, 0))) +
  scale_x_discrete(expand = expansion(mult = c(0, 0))) +
  scale_fill_manual(values = c("Yes" = "red", "No" = "white")) +
  theme_bw() +
  labs(x = "", y = "") +
  theme(axis.text = element_text(
    angle = 45,
    hjust = 1,
    vjust = 1
  ),
  axis.ticks.y = element_blank(),
  plot.margin = unit(c(0, 0, 0, 0), "pt"),
  panel.grid.minor = element_blank())

plot2 <-
  total_sample_info %>% 
    dplyr::filter(value == "Yes") %>% 
  dplyr::group_by(class) %>% 
  dplyr::summarise(n = n()) %>% 
    dplyr::ungroup() %>% 
  ggplot(aes(date, class)) +
    geom_bar(aes(x = n, y = class), stat = "identity", color = "black") +
    scale_x_continuous(expand = expansion(mult = c(0, 0))) +
    scale_y_discrete(expand = expansion(mult = c(0, 0))) +
    scale_fill_manual(values = c("Yes" = "red", "No" = "white")) +
    theme_classic() +
    labs(x = "Sample number", y = "") +
    theme(axis.text.x = element_text(
      size = 12
    ),
    axis.text.y = element_blank(),
    axis.title = element_text(size = 13),
    axis.ticks.y = element_blank(),
    plot.margin = unit(c(0, 0, 0, 0), "pt"),
    panel.grid.minor = element_blank())

plot3 <-
  total_sample_info  %>% 
  dplyr::filter(value == "Yes") %>% 
  dplyr::group_by(date, class) %>% 
  dplyr::summarise(n = n()) %>% 
  dplyr::ungroup() %>% 
  ggplot(aes(date, n)) +
  geom_bar(aes(x = date, y = n, fill = class), 
           stat = "identity", color = "black", show.legend = FALSE) +
  guides(fill = guide_legend(nrow = 1)) +
  scale_x_discrete(expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(expand = expansion(mult = c(0, 0))) +
  scale_fill_manual(values = value) +
  guides(fill = guide_legend(nrow = 2)) +
  theme_classic() +
  labs(x = "", y = "Sample number") +
  theme(axis.text.y = element_text(
    size = 12
  ),
  axis.text.x = element_blank(),
  axis.title = element_text(size = 13),
  axis.ticks.x = element_blank(),
  plot.margin = unit(c(0, 0, 0, 0), "pt"),
  panel.grid.minor = element_blank(),
  legend.position = "top")

plot <-
  {
    plot3 + plot2 + plot_layout(ncol = 2, widths = c(5, 1))
  } -
  {
    plot1 + plot2 + plot_layout(ncol = 2, widths = c(5, 1))
  } +
  plot_layout(ncol = 1, heights = c(1, 5))

plot
```




